---
title: "Lab 9: Bab"
author: "Cameron White"
format: 
  html:
    code-fold: true
    code-summary: "show code"
editor: visual
execute: 
  output: false
---

#### Setup

```{r}
library(tidyverse)


bab_F_A <- read_csv(
  here::here(
    "supporting_artifacts", "Labs", "Lab 9", "StateNames_A.csv")) |> 
  rename(
    "Sex" = "Gender") |> 
  filter(
    Sex == 'F')

```

#### Step 3: Time-Series Visualization. Allisons Across Eons

```{r}
alli_sum <- bab_F_A |>
  filter(
    Name == 'Allison') |> 
  group_by(
    Year, Name) |> 
  summarise(
    Count = sum(Count))


alli_sum |> 
  ggplot(mapping = aes(
    x = Year,
    y = Count)) +
  geom_line(
    color = "tomato",
    size = 0.9) +
  geom_point(
    color = "darkblue"
  ) +
  labs(
    x = "",
    y = "") +
  ggtitle("United States: # of female children born named Allison (1997-2014)") + 
  theme(plot.title.position = "plot") +
  scale_y_continuous(
    breaks = seq(5500, 7500, 1000)) +
  scale_x_continuous(
    breaks = seq(1994, 2016, 4))

```

#### Step 4: Modeling the Number of Allisons

###### Linear Model: Allison by Year function

```{r}
# Calculate the linear model
lm_alli_sum <- alli_sum |> 
  lm(
    Count ~ Year, 
    data = _)

# Visualize the linear model
alli_sum |> 
  ggplot(mapping = aes(
    x = Year,
    y = Count)) +
  geom_line(
    color = "tomato",
    size = 0.9) +
  geom_point(
    color = "darkblue"
  ) +
  labs(
    x = "",
    y = "") +
  ggtitle(
    "United States: # of female children born named Allison (1997-2014)") + 
  theme(plot.title.position = "plot") +
  scale_y_continuous(
    breaks = seq(5500, 7500, 1000)) +
  scale_x_continuous(
    breaks = seq(1994, 2016, 4)) +
  geom_smooth(
    method = "lm",
    se = F)
  
```

###### The linear model regression equation is:

###### *y = 209689 - (101.5 /* x)\*

#### Linear Model: Plotting the Residuals

```{r}
broom::augment(lm_alli_sum) |> 
  ggplot(mapping = aes(
    x = .fitted,
    y = .resid
  )) +
  geom_point() +
  geom_smooth(se = F)

```

###### There are too few values to make a reasonable assumption about the data. But if I were forced to, it seems that the model overestimates values between \~6000 and and \~6700. There are a few other values that are far underestimated by the linear model, but they don't seem to be part of a larger pattern.

###### I conclude from the model that the name "Allison" is in somewhat severe decline, though counts refuse to go below \~5500 Allisons per year. From its high in 1998, it has declined in popularity by about 25%. It is difficult to say if the name "Allison" is truly lame-o without a fuller examination of counts against other names.

######But, according to this analysis, it is not as attractive to parents as it once was. An important follow-up question may be: "Are we using an appropriate metric for 'cool'?"

#### Part 5: Spelling by State

#### Allan-metry

```{r}
bab_M_A <- read_csv(
  here::here(
    "Labs", "Lab 9", "StateNames_A.csv")) |> 
  rename(
    "Sex" = "Gender") |>
  filter(
    Sex == 'M',
    Name == "Allan" | 
    Name == "Alan" | 
    Name == "Allen")

unique(bab_M_A$Name)
```

#### Allan-patric Inspection

```{r}
# Filter and Pivot
a <- bab_M_A |> 
  filter(
    Year == 2000,
    State == 'CA'|
    State == 'PA') |> 
  pivot_wider(
    names_from = Name,
    values_from = Count)
  select(
    Alan:Allan)
  
  |>
  apply(1, FUN = sum) |> 
    cbind
  
?apply
  
a <- a |>
  filter(State == 'CA') |> 
  mutate("Alan_pct" = (Alan/(sum(Alan, Allen, Allan))))
 
a <- a |> 
  mutate("Alan_pct" = (Alan/(sum(Alan, Allen, Allan))))
 
```
